---
title: "Check Data for TS"
output: html_notebook
---

Aim: check quality of chosen variables for panel analysis (time series: TS).

The notebook takes some time to run, because of the *facet_wrap* by country, so if you don't want to re-run code, just read the html!

```{r settings}

library(tidyverse)
library(naniar)


# variables of interest:
vars = c('vdem_edcomp_thick', 'vdem_partip', 'vdem_corr', 'vdem_egal', 'fh_cl', 'fh_pr')

# data set:
ts <- readRDS("../../../Data/Data for Modelling/LEGTER_ts.rds")

# select only columns of interest:
ts <- ts %>%
  select(year, 
         consolidated_country,
         one_of(vars)
         )

# glimpse(ts)
ts
```

It seems that, by just quickly looking at the data, the _Varieties of Democracy (V-Dem)_ variables have a higher granularity and vary over the years, whereas the _Freedom of House (fh)_ vary less over the years. 


## Check NAs

Let's see if we have many NAs, and how they are distributued.  

We have all years and countries and we miss a few combinations for the V-Dem and FH variables.  


```{r vis_NA}

# tmp <- na.omit(ts)

vis_miss(ts)

```

Which countries are missing data? First we can explore all observations with missing values: 

```{r obs_NA}

ts[!complete.cases(ts), ]

```

And those are the countries with missing values. We have a few large countries in there (France, Pakistan, Malaysia, ...).

**==> TO DO: check if correct that no V-Dem and FH data for France and Pakistan**

```{r ctry _NA}

ts[!complete.cases(ts), ]$consolidated_country %>% unique()

```

## Check Distributions

Looking at the distribution of the variables of interest:

```{r distr_fh}
ts %>%
  select(fh_cl, fh_pr) %>%
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(stat = "count") +
  theme_minimal() + 
  ggtitle("Distributions for Freedom's House Index variables")

```


```{r distr_vdem_bin}
ts %>%
  select(vdem_edcomp_thick, vdem_partip, vdem_corr, vdem_egal) %>%
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(stat = "bin", bins = 20) +
  theme_minimal() + 
  ggtitle("Distributions for V-Dem's variables, 20 bins")
```


```{r distr_vdem_density}
ts %>%
  select(vdem_edcomp_thick, vdem_partip, vdem_corr, vdem_egal) %>%
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_density() +
  theme_minimal() + 
  ggtitle("Distributions for V-Dem's variables, density")
```



## Check Trends

We want to look at the variability over time of the variables.

### vdem_edcomp_thick

```{r trends_vdem_edcomp_thick, fig.width = 10, fig.height = 15}
ts %>%
  ggplot(aes(x=year, y=vdem_edcomp_thick)) +
  geom_line() +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("vdem_edcomp_thick: trends by country") +
  facet_wrap(~ consolidated_country)

```

### vdem_partip

```{r vdem_partip, fig.width = 10, fig.height = 15}
ts %>%
  ggplot(aes(x=year, y=vdem_partip)) +
  geom_line() +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("vdem_partip: trends by country") +
  facet_wrap(~ consolidated_country)
```


### vdem_corr

```{r vdem_corr, fig.width = 10, fig.height = 15}
ts %>%
  ggplot(aes(x=year, y=vdem_corr)) +
  geom_line() +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("vdem_corr: trends by country") +
  facet_wrap(~ consolidated_country)
```

### vdem_egal

```{r vdem_egal, fig.width = 10, fig.height = 15}
ts %>%
  ggplot(aes(x=year, y=vdem_egal)) +
  geom_line() +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("vdem_egal: trends by country") +
  facet_wrap(~ consolidated_country)
```


### fh_cl

```{r fh_cl, fig.width = 10, fig.height = 15}
ts %>%
  ggplot(aes(x=year, y=fh_cl)) +
  geom_line() +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("fh_cl: trends by country") +
  facet_wrap(~ consolidated_country)
```

### fh_pr

```{r fh_pr, fig.width = 10, fig.height = 15}
ts %>%
  ggplot(aes(x=year, y=fh_pr)) +
  geom_line() +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("fh_pr: trends by country") +
  facet_wrap(~ consolidated_country)
```




